Stochastic Variance-Reduced Policy Gradient

### 2018-02-09 svrpg.tex
Fit in 8 pages

\begin{figure*}[t]
    \begin{minipage}[t]{0.32\textwidth}
	\includegraphics[width=.98\textwidth]{images/cart_pole_SVRPG_vs_GPOMDP_tex.pdf}
	\caption{On-line performance over sampled trajectories of SVRPG vs G(PO)MDP in the Cart-pole task, with 90\% confidence intervals.}
	\label{fig:cartpole}
    \end{minipage}
    \hfill
    \begin{minipage}[t]{0.32\textwidth}
	\includegraphics[width=0.98\textwidth]{swimmer_self_normalized_SVRPG_vs_SVRPG_tex.pdf}
	\caption{On-line performance over sampled trajectories of self-Normalized SVRPG vs unbiased SVRPG in the Swimmer task, with 90\% confidence intervals.}
	\label{fig:swimmertwo}
    \end{minipage}
    \hfill
    \begin{minipage}[t]{.32\textwidth}
	\includegraphics[width=0.98\textwidth]{swimmer_self_normalized_SVRPG_vs_GPOMDP_tex.pdf}
	\caption{On-line performance over sampled trajectories of self-Normalized SVRPG vs G(PO)MDP in the Swimmer task, with 90\% confidence intervals}
	\label{fig:swimmerone}
    \end{minipage}
\end{figure*}

\begin{table*}[t]
	\caption{
		\label{tab:1}
		Comparison between G(PO)MDP and SVRPG on different tasks. Average return over the whole learning process is reported, together with normalized standard deviation and 90\% confidence intervals. Results denoted with $\dagger$ are taken from \cite{duan2016benchmarking}.}
	\centering
	\begin{tabular}{@{}lccccccc@{}} 
		\toprule
		\phantom{abc} & \multicolumn{3}{c}{G(PO)MDP} & \phantom{abc} & \multicolumn{3}{c}{SVRPG} \\
		\cmidrule{2-4} \cmidrule{6-8}
		\phantom{abc} & Avg Return & Norm. Std & C.I. & \phantom{abc}
		& Avg Return & Norm. Std & C.I.
		\\\cmidrule{2-8}
		Cart-pole & $618.49$ & $21.09$ & $[579.82, 657.16]$ & \phantom{abc} 
		& $736.31$ & $16.51$ & $[706.04, 766.57]$ \\
		Swimmer & $39.86$ & $3.84$ & $[32.82, 46.89]$ & \phantom{abc} & $49.60$ & $3.27$ & $[43.60, 55.60]$\\
		Half-cheetah & $1183.1^\dagger$ & $69.20^\dagger$ & $[1035.58, 1330.62]^\dagger$ & \phantom{abc} & $1480.95$ & $73.21$ & $[1324.87, 1637.03]$\\
		\bottomrule
	\end{tabular}
\end{table*}

@@

\vspace{-0.05in}
\section{Conclusion}
\vspace{-0.05in}
In this paper, we introduced SVRPG, a variant of SVRG designed explicitly for RL problems.
The control problem considered in the paper has a series of difficulties that are not common in SL.
Among them, non-concavity and approximate estimates of the FG have been analysed independently in SL~\citep[\eg][]{allen2016variance,reddi2016stochastic,harikandeh2015stopwasting} but never combined.
Nevertheless, the main issue in RL is the non-stationarity of the sampling process since the distribution underlying the objective function is policy-dependent.
We have shown that by exploiting importance weighting techniques, it is possible to overcome this issue and preserve the unbiasedness of the corrected gradient.
We have additionally shown that, under mild assumptions that are often verified in RL applications, it is possible to derive convergence guarantees for SVRPG.
%{\color{red}SVRG has been previously used in RL for policy evaluation~\citep{du2017svrgpe} and as alternative optimization method for TRPO~\citep{xu2017svrgtrpo}.
%The first case is much more similar to the SL case since the problem is stationary and may be made concave, removing several issues we have faced here.
%In the second case, it was used as a better optimization technique to solve the per-iteration problem imposed by TRPO.
%}
%\todomatout{Repetition of Section 6?}
Finally, we have empirically shown that practical variants of the theoretical SVRPG version can outperform classical actor-only approaches on benchmark tasks.
Preliminary results support the effectiveness of SVRPG also with a commonly used baseline for the policy gradient.
Despite that, we believe that it will be possible to derive a baseline designed explicitly for SVRPG to exploit the RL structure and the SVRG idea jointly. Future work should also focus on making batch sizes $N$ and $B$ adaptive, as suggested in \cite{papini2017adaptive}.

>>>>

\begin{figure*}[t]
    \begin{minipage}[t]{0.32\textwidth}
	\includegraphics[width=.98\textwidth]{images/cart_pole_SVRPG_vs_GPOMDP_tex.pdf}
	\vspace{-0.1in}
	\caption{On-line performance over sampled trajectories of SVRPG vs G(PO)MDP in the Cart-pole task, with 90\% confidence intervals.}
	\label{fig:cartpole}
    \end{minipage}
    \hfill
    \begin{minipage}[t]{0.32\textwidth}
	\includegraphics[width=0.98\textwidth]{swimmer_self_normalized_SVRPG_vs_SVRPG_tex.pdf}
	\vspace{-0.1in}
	\caption{On-line performance over sampled trajectories of self-Normalized SVRPG vs unbiased SVRPG in the Swimmer task, with 90\% confidence intervals.}
	\label{fig:swimmertwo}
    \end{minipage}
    \hfill
    \begin{minipage}[t]{.32\textwidth}
	\includegraphics[width=0.98\textwidth]{swimmer_self_normalized_SVRPG_vs_GPOMDP_tex.pdf}
	\vspace{-0.1in}
	\caption{On-line performance over sampled trajectories of self-Normalized SVRPG vs G(PO)MDP in the Swimmer task, with 90\% confidence intervals}
	\label{fig:swimmerone}
    \end{minipage}
    \vspace{-0.15in}
\end{figure*}

\begin{table*}[t]
	\caption{
		\label{tab:1}
		Comparison between G(PO)MDP and SVRPG on different tasks. Average return over the whole learning process is reported, together with normalized standard deviation and 90\% confidence intervals. Results denoted with $\dagger$ are taken from \cite{duan2016benchmarking}.}
	\centering
	\begin{tabular}{@{}lccccccc@{}} 
		\toprule
		\phantom{abc} & \multicolumn{3}{c}{G(PO)MDP} & \phantom{abc} & \multicolumn{3}{c}{SVRPG} \\
		\cmidrule{2-4} \cmidrule{6-8}
		\phantom{abc} & Avg Return & Norm. Std & C.I. & \phantom{abc}
		& Avg Return & Norm. Std & C.I.
		\\\cmidrule{2-8}
		Cart-pole & $618.49$ & $21.09$ & $[579.82, 657.16]$ & \phantom{abc} 
		& $736.31$ & $16.51$ & $[706.04, 766.57]$ \\
		Swimmer & $39.86$ & $3.84$ & $[32.82, 46.89]$ & \phantom{abc} & $49.60$ & $3.27$ & $[43.60, 55.60]$\\
		Half-cheetah & $1183.1^\dagger$ & $69.20^\dagger$ & $[1035.58, 1330.62]^\dagger$ & \phantom{abc} & $1480.95$ & $73.21$ & $[1324.87, 1637.03]$\\
		\bottomrule
	\end{tabular}
    \vspace{-0.05in}
\end{table*}

@@

\vspace{-0.05in}
\section{Conclusion}
\vspace{-0.05in}
In this paper, we introduced SVRPG, a variant of SVRG designed explicitly for RL problems.
The control problem considered in the paper has a series of difficulties that are not common in SL.
Among them, non-concavity and approximate estimates of the FG have been analysed independently in SL~\citep[\eg][]{allen2016variance,reddi2016stochastic,harikandeh2015stopwasting} but never combined.
Nevertheless, the main issue in RL is the non-stationarity of the sampling process since the distribution underlying the objective function is policy-dependent.
We have shown that by exploiting importance weighting techniques, it is possible to overcome this issue and preserve the unbiasedness of the corrected gradient.
We have additionally shown that, under mild assumptions that are often verified in RL applications, it is possible to derive convergence guarantees for SVRPG.
%{\color{red}SVRG has been previously used in RL for policy evaluation~\citep{du2017svrgpe} and as alternative optimization method for TRPO~\citep{xu2017svrgtrpo}.
%The first case is much more similar to the SL case since the problem is stationary and may be made concave, removing several issues we have faced here.
%In the second case, it was used as a better optimization technique to solve the per-iteration problem imposed by TRPO.
%}
%\todomatout{Repetition of Section 6?}
Finally, we have empirically shown that practical variants of the theoretical SVRPG version can outperform classical actor-only approaches on benchmark tasks.
Preliminary results support the effectiveness of SVRPG also with a commonly used baseline for the policy gradient.
Despite that, we believe that it will be possible to derive a baseline designed explicitly for SVRPG to exploit the RL structure and the SVRG idea jointly.
Another possible improvement would be to employ the natural gradient~\cite{kakade2002natural} to better control the effects of parameter updates on the variance of importance weights. Future work should also focus on making batch sizes $N$ and $B$ adaptive, as suggested in~\cite{papini2017adaptive}.
